---
title: "Cross Sectional Profile Lines"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cross Sectional Profile Lines}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Cross-section profiles are valuable tools in hydrology for understanding, analyzing, and managing water resources and drainage systems. We will use a stream center line GIS file to generate perpendicular cross-sectional profiles.


```{r setup}
# Load or install the streamgis package
library(streamgis)

# Import a simple stream cetner line
# center_line <- st_read("./path/to/my/file.gpkg", layer = "layer name")

# or use default provided for tutorial
fname <- system.file("extdata", "center_line.gpkg", package="streamgis")
center_line <- sf::st_read(fname)
plot(sf::st_geometry(center_line))

```


### Sample points along line

Sample points along line or supply custom sample locations. Be sure to set the correct EPS G code for a local UTM zone. Set the point spacing to units of meters. `points_on_line` will return a `sf` spatial points object with line segment ID, point ID and cumulative distance. 


```{r pol}

points <- suppressWarnings({ points_on_line(center_line,
                                         point_spacing = 100,
                                         epsg = 26910) })
head(points, 3)

plot(sf::st_geometry(center_line[center_line$id == 1, ]))
plot(sf::st_geometry(points[points$l_id == 1, ]), add = TRUE, col = "red")


```

### Create cross-sectional profiles

Next we can create cross-sectional profile lines at each of the sample points along the stream centre line. Cross-sectional profile lines will be drawn perpendicular to the stream centre line. Additional information such as distance and bearings are in the data file as attribute fields.


```{r cls}

csl <- cross_section_lines(
  center_line = center_line,
  points = points,
  cross_profile_length = 400,
  epsg = 26910)

head(csl, 3)

plot(sf::st_geometry(center_line[center_line$id == 1, ]))
plot(sf::st_geometry(points[points$l_id == 1, ]), add = TRUE, col = "red")
plot(sf::st_geometry(csl[csl$l_id == 1, ]), add = TRUE, col = "blue")

# sf::st_write(points, dsn = "points.gpkg", delete_dsn = TRUE)
# sf::st_write(csl, dsn = "profiles.gpkg", delete_dsn = TRUE)




```





